API design for machine learning software: experiences from the scikit-learn project

### 2013-07-30 paper.tex
explain why estimators serve as models
Third reviewer wanted more extensive comparison with alternative
approaches.

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
As such, it always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.
The choice to let a single object serve dual purpose as estimator and model
simplifies the \sklearn class hierarchy
compared to an approach were estimators and models are separated.

>>>>

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
It always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.

The choice to let a single object serve dual purpose as estimator and model
simplifies the \sklearn API
compared to an approach where estimators and models are separated,
forming two parallel class hierarchies.
While from a mathematical point of view,
models are more similar to one another than algorithms,
we argue that this is less true in practical machine learning.
As an example, consider the case of linear models,
described mathematically by a weight matrix $w$ and an intercept vector $b$,
suggesting a class \texttt{LinearModel}
with such a matrix and vector as attributes.
However, to use such a model in practice,
one first needs to know whether the task at hand is
regression, $y = w^\top x + b$,
or classification, $y = \arg \max_k[w_{:,k}^\top x + b_k]$,
calling for two classes.
Then, the various linear probabilistic classifiers must be distuingished:
logistic regression, with $P(y \mid x) = \sigma(w^\top x + b)$,
multinomial naïve Bayes, with $\log P(y \mid x) = w^\top x + b$,
and the winnow probability estimator \citep{zhang2002}, with
$P(y \mid x) = \frac{1}{2} (1 + \mathrm{median} \{-1, w^\top x + b, 1 \})$,
would all need either distinct classes,
or a common class that remembers the training regime.
Furthermore, separating estimators and models
would make online/out-of-core learning
(not further discussed in the present paper) more complicated.
